Respiration Related Death Cause show Signficant Relation to Increasing 

### 2024-03-14 paper/paper.qmd
more updates

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - First author
  - Another author
thanks: "Code and data are available at: LINK."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

@@

# Plot the regression graph
ggplot(data, aes(x = total_deaths, y = provincial_average)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +  # Add linear regression line without confidence interval
  labs(x = "Total Deaths", y = "Provincial Average", title = "Regression Graph") +
  theme_minimal()

# Plot the regression graph
ggplot(data_peak, aes(x = total_deaths, y = provincial_average)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +  # Add linear regression line without confidence interval
  labs(x = "Total Deaths", y = "Provincial Average", title = "Regression Graph") +
  theme_minimal()

# Plot the regression graph
ggplot(data_peak, aes(x = provincial_average, y = total_deaths)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +  # Add linear regression line without confidence interval
  labs(x = "Provincial Average", y = "Total Deaths", title = "Regression Graph") +
  theme_minimal()

@@

# Plot the regression graph for all lungs causes
ggplot(data_lung, aes(x = provincial_average, y = total_deaths)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +  # Add linear regression line without confidence interval
  labs(x = "Provincial Average", y = "Total Deaths", title = "Regression Lung") +
  theme_minimal()

# Plot the regression graph for all heart causes
ggplot(data, aes(x = provincial_average, y = total_deaths)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +  # Add linear regression line without confidence interval
  labs(x = "Provincial Average", y = "Total Deaths", title = "Regression Heart") +
  theme_minimal()

@@

# analysis_data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))
# 
# analysis_data |> 
#   ggplot(aes(x = width, y = length)) +
#   geom_point(alpha = 0.8) +
#   theme_minimal() +
#   labs(x = "Wing width (mm)",
#        y = "Wing length (mm)")
```

@@

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
# 
# library(rstanarm)
# 
# first_model <-
#   readRDS(file = here::here("models/first_model.rds"))
# 
# poisson_model <-
#   readRDS(file = here::here("models/poisson_model.rds"))
# 
# causes_model <-
#   readRDS(file = here::here("models/causes_model.rds"))
# 
# p_average_model <-
#   readRDS(file = here::here("models/p_average_model.rds"))
#

@@

# Assuming you've already read your models as before
neg_binom_model <- readRDS(file = here::here("models/neg_binom_model.rds"))
stan_neg_binom_model <- readRDS(file = here::here("models/stan_neg_binom_model.rds"))
stan_poisson_model <-readRDS(file = here::here("models/stan_poisson_model.rds"))
peak_neg_binom_model <- readRDS(file = here::here("models/peak_neg_binom_model.rds"))
peak_poisson_model <-readRDS(file = here::here("models/peak_poisson_model.rds"))
first_model <-
  readRDS(file = here::here("models/first_model.rds"))

@@

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false
library(modelsummary)
# modelsummary::modelsummary(
#   list(
#     "First model" = first_model
#   ),
#   #statistic = "mad",
#   fmt = 2
# )
# 
# modelsummary::modelsummary(
#   list(
#     "First model" = poisson_model
#   ),
#   #statistic = "mad",
#   fmt = 2
# )
# 
# modelsummary::modelsummary(
#   list(
#     "First model" = causes_model
#   ),
#   #statistic = "mad",
#   fmt = 2
# )
# 
# modelsummary::modelsummary(
#   list(
#     "First model" = p_average_model
#   ),
#   #statistic = "mad",
#   fmt = 2
# )
#

@@

pp_check(peak_poisson_model) +
  theme(legend.position = "bottom")

pp_stan_neg_binom_model <- pp_check(peak_poisson_model)
pp_stan_neg_binom_model + theme(legend.position = "bottom")

pp_check(peak_neg_binom_model) +
  theme(legend.position = "bottom")

pp_stan_neg_binom_model <- pp_check(peak_neg_binom_model)
pp_stan_neg_binom_model + theme(legend.position = "bottom")

@@

plot(first_model, "rhat")
```

>>>>

---
title: "Respiration Related Death Cause show Signficant Relation to Increasing "
subtitle: "My subtitle if needed"
author: 
  - Vanshika Vanshika
  - Shivank Goel
  - Navya Hooda
thanks: "Code and data are available at: https://github.com/shivankgoel003/Mortality-in-Alberta."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

@@

@@

# Plot the regression graph for all lungs causes
ggplot(data_lung, aes(x = provincial_average, y = total_deaths)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +  # Add linear regression line without confidence interval
  labs(x = "Provincial Average", y = "Total Deaths", title = "Regression Lung") +
  theme_minimal()
```

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false
# Plot the regression graph for all heart causes
ggplot(data, aes(x = provincial_average, y = total_deaths)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +  # Add linear regression line without confidence interval
  labs(x = "Provincial Average", y = "Total Deaths", title = "Regression Heart") +
  theme_minimal()

@@

# Assuming your data frame is named 'cleaned_avg_PM25_conc' and is already filtered and renamed.
# Convert factors to character to avoid issues in ggplot
data[] <- lapply(data, function(x) if(is.factor(x)) as.character(x) else x)
# Creating the plot with ggplot
p <- ggplot(data, aes(x = year)) +
  geom_line(aes(y = provincial_average, colour = "Provincial Average")) +
  geom_line(aes(y = x10th_percentile, colour = "10th Percentile")) +
  geom_line(aes(y = x90th_percentile, colour = "90th Percentile")) +
  labs(title = "Annual Trends of PM2.5 Concentrations in Alberta: Provincial Averages and Percentile Ranges (2012-2021)",
       x = "Years", y = "PM2.5 Concentration (μg/m³)") +
  scale_colour_manual(values = c("Provincial Average" = "blue", "10th Percentile" = "green", "90th Percentile" = "red")) +
  theme_minimal()

# Display the plot
p

library(MASS)

air_quality_data <- air_quality_data[air_quality_data$air_quality_health_index != "10+",]
air_quality_data$air_quality_health_index <- as.numeric(air_quality_data$air_quality_health_index)

# Calculate the average Air Quality Health Index per year
average_aqi_by_year <- air_quality_data %>%
  group_by(year) %>%
  summarize(average_aqi = mean(air_quality_health_index))

combined_data <- merge(deaths_data, average_aqi_by_year, by = "year")

# Plot scatter plot of mortality rates vs. air quality health index
ggplot(combined_data, aes(x = average_aqi, y = total_deaths)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_wrap(~cause) +
  labs(title = "Scatter Plot of Mortality Rates vs. Air Quality Health Index",
       x = "Average Air Quality Health Index",
       y = "Total Deaths")

# Fit the model
model <- glm.nb(total_deaths ~ average_aqi + factor(year) + factor(cause), data = combined_data)

# Model Diagnostics
par(mfrow=c(2,2))
plot(model)

# Add Predictions to the combined data
# Note: predict() function might throw an error if there are new factor levels in the test set that were not present in the training set
# To prevent this, the newdata argument should only include levels that were present in the model fitting
combined_data$predicted_deaths <- predict(model, newdata = combined_data, type = "response")

# Plot Observed vs. Predicted Deaths
ggplot(combined_data, aes(x = total_deaths, y = predicted_deaths)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "red") +
  labs(x = "Observed Deaths", y = "Predicted Deaths") +
  theme_minimal()

# Effect of AQHI on Predicted Deaths
ggplot(combined_data, aes(x = average_aqi, y = predicted_deaths, color = factor(year))) +
  geom_point() +
  geom_line(aes(group = factor(year))) + # Adding lines to connect points by year
  labs(x = "Average Air Quality Health Index", y = "Predicted Deaths") +
  theme_minimal() +
  facet_wrap(~cause)

# Merge datasets on 'year'
combined_data <- merge(deaths_data, average_aqi_by_year, by = "year")

# Plot scatter plot of mortality rates vs. air quality health index
ggplot(combined_data, aes(x = average_aqi, y = total_deaths)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_wrap(~cause) +
  labs(title = "Scatter Plot of Mortality Rates vs. Air Quality Health Index",
       x = "Average Air Quality Health Index",
       y = "Total Deaths")

# Summarize total deaths for each disease by year
total_deaths_by_year <- deaths_data %>%
  group_by(year, cause) %>%
  summarize(total_deaths = sum(total_deaths))

# Plot for mortality data
ggplot(total_deaths_by_year, aes(x = as.factor(year), y = total_deaths, group = cause, color = cause)) +
  geom_line() +
  labs(title = "Total Deaths by Year for Each Disease",
       x = "Year",
       y = "Total Deaths")

# Ensure the 'year' column is in numeric format for proper line plotting
average_aqi_by_year$year <- as.numeric(as.character(average_aqi_by_year$year))

# Plotting with ggplot2
ggplot(average_aqi_by_year, aes(x = year, y = average_aqi)) +
  geom_line() +  # for lines
  geom_point() + # for points
  labs(title = "Average Air Quality Health Index by Year",
       x = "Year",
       y = "Average Air Quality Index") +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 10)) # Optional: Adjust x-axis breaks

ggplot(average_aqi_by_year, aes(x = as.factor(year), y = average_aqi, fill = as.factor(year))) +
  geom_bar(stat = "identity") +
  labs(title = "Average Air Quality Health Index by Year",
       x = "Year",
       y = "Average AQHI") +
  theme_minimal()

@@

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

@@

# Assuming you've already read your models as before
neg_binom_model <- readRDS(file = here::here("models/neg_binom_model.rds"))
stan_neg_binom_model <- readRDS(file = here::here("models/stan_neg_binom_model.rds"))
stan_poisson_model <-readRDS(file = here::here("models/stan_poisson_model.rds"))

heart_model <- readRDS(file = here::here("models/heart_model.rds"))
lung_model <-readRDS(file = here::here("models/lung_model.rds"))

first_model <-
  readRDS(file = here::here("models/first_model.rds"))

@@

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Model of "
#| warning: false
library(modelsummary)

modelsummary::modelsummary(
  list(
    "heart causes model" = heart_model
  ),
  #statistic = "mad",
  fmt = 2
)

modelsummary::modelsummary(
  list(
    "Lung Causes model" = lung_model
  ),
  #statistic = "mad",
  fmt = 2
)

@@

pp_check(heart_model) +
  theme(legend.position = "bottom")

pp_check(lung_model) +
  theme(legend.position = "bottom")

@@

plot(first_model, "rhat")

plot(stan_neg_binom_model, "trace")

plot(stan_neg_binom_model, "rhat")
```
